# Loading and preprocessing the data.  


###**1. Load the data.**
```{r}
library(data.table)
activity <- read.csv('activity.csv')
```


###**2. Process the data.**
```{r}
activity$date <- as.Date(as.character(activity$date))
head(activity)
```



# What is the mean total number of steps taken per day?
```{r Data_per_day}
daily_activity <- split(activity, activity$date)
days <- length(daily_activity)
daily_steps <- NULL
daily_mean <- NULL
daily_median <- NULL
for (i in 1:days){
        daily_steps <- c(daily_steps, sum(daily_activity[[i]]$steps,na.rm = TRUE))
}

daily <- data.frame('date'= unique(activity$date),
                    'daily_steps'=  daily_steps)
head(daily)
```


###**1. Histogram of the total number steps per day.**

```{r}
hist(daily$daily_steps)
```


###**2. Mean and median of the total number of steps taken each day.**

```{r}
mean_total_steps <- mean(daily$daily_steps)
median_total_steps <- median(daily$daily_steps)
```

**Mean** is `r mean_total_steps` and **median** is `r median_total_steps`.

# What is the average daily activity pattern?

###**1. Panel plot representing (y-axis) the average number of steps taken, averaged across all days, in function of the 5-minute interval (x-axis).**

```{r}
x <- daily_activity[[1]]$interval

y <- NULL
number_of_intervals <- nrow(daily_activity[[1]])
interval_activity <- split(activity, activity$interval)
for (i in 1:number_of_intervals){
        y <- c(y, mean(interval_activity[[i]]$steps, na.rm = TRUE))
}
plot(x,y, type='l', xlab = 'interval')
title('Average Daily Activity')
interval_max <- x[which.max(y)]
```

The **5-minute interval, on average across all the days in the dataset, that contains the maximum number of steps** is `r interval_max`.




# Imputing missing values.

###**1. Total number of missing values in the dataset.**

```{r}
length(activity$steps[is.na(activity$steps)])
```


###**2. Filling in missing values.** 

We decide to replace each missing value in a certain interval with the **mean of the values in this interval**.

```{r}
filling_values <- y

filled_in_interval_activity <- interval_activity
for (i in 1:number_of_intervals){
        index <- which(is.na(interval_activity[[i]]$steps))
        filled_in_interval_activity[[i]]$steps[index] <- rep(filling_values[i], length(index))
}
```

###**3. Here is the filled in dataset.**

```{r}
filled_in_activity <- rbindlist(filled_in_interval_activity)
index <- with(filled_in_activity, order(date, interval))
filled_in_activity <- as.data.frame(filled_in_activity[index, ])
head(filled_in_activity)
```


###**4. Histogram, mean and median of the total of the total number of steps per day for the filled in dataset.**

```{r}
filled_in_daily_activity <- split(filled_in_activity, filled_in_activity$date)
daily_steps <- NULL
daily_mean <- NULL
daily_median <- NULL
for (i in 1:days){
        daily_steps <- c(daily_steps, 
                                   sum(filled_in_daily_activity[[i]]$steps))
}

filled_in_daily <- data.frame('date'= unique(activity$date),
                              'daily_steps' = daily_steps)

hist(filled_in_daily$daily_steps)
filled_in_mean_total_steps <- mean(filled_in_daily$daily_steps)
filled_in_median_total_steps <- median(filled_in_daily$daily_steps)
diff_mean <- abs(mean_total_steps-filled_in_mean_total_steps)
diff_median <- abs(median_total_steps-filled_in_median_total_steps)
```

**Mean** is `r filled_in_mean_total_steps` and **median** is `r filled_in_median_total_steps`.

###**Observations:** 

Regarding the **mean** and **median** values:  

The **initial and second mean values** are respectively `r mean_total_steps` and `r filled_in_mean_total_steps`, and thus differ from `r diff_mean`, which is significant.  

On the hand, the **initial and second median values** are respectively `r median_total_steps` and `r filled_in_median_total_steps`, and thus don't differ much (`r diff_median`).    

Besides, one can see that filling in the missing values does affect 
the initial histogram. Indeed, the obtained frequency of the modified dataset is clearly better **normally** distributed (bell curve). Also, note that the **mean and median values** (`r filled_in_mean_total_steps` and `r filled_in_median_total_steps` resp.) for the modified dataset don't differ much unlike in the first place (`r mean_total_steps` and `r median_total_steps` resp.), as it is the case for **normally** distributed frequencies. 


# Are there differences in activity patterns between weekdays and weekends?

```{r}
new_activity <- filled_in_activity
````


###**1. Creating a new factor variable in the dataset 'new_activity'.**

```{r}
day_type <- function(dates){
        weekday <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
        weekend <- c('Saturday', 'Sunday')
        days <- weekdays(dates)
        results <- NULL
        for (day in days){
                check <- day == weekday
                if (any(check)){
                        results <- c(results,'weekday')
                }else{
                        results <- c(results,'weekend') 
                }
        }
        factor(results, levels = c('weekday', 'weekend'))
}

new_activity$'day_type' <- day_type(new_activity$date)
head(new_activity)
```



###**2. Panel plot representing (y-axis) the average number of steps taken (averaged across all weekday days or weekend days) in function of the 5-minute interval (x-axis).**

```{r}
x <- unique(new_activity$interval)

y_weekday <- NULL
y_weekend <- NULL
day_type_split <- split(new_activity, new_activity$day_type)
new_activity_weekday <- day_type_split$'weekday'
new_activity_weekend <- day_type_split$'weekend'
interval_new_activity_weekday <- split(new_activity_weekday, new_activity_weekday$interval)
interval_new_activity_weekend <- split(new_activity_weekend, new_activity_weekend$interval)
for (i in 1:number_of_intervals){
        y_weekday <- c(y_weekday, mean(interval_new_activity_weekday[[i]]$steps, na.rm = TRUE))
        y_weekend <- c(y_weekend, mean(interval_new_activity_weekend[[i]]$steps, na.rm = TRUE))
}

par(mfrow=c(2,1))
plot(x, y_weekday, type='l', xlab = 'interval')
title('Weekday')
plot(x, y_weekend, type='l', xlab = 'interval')
title('Weekend')
```
